---
title: "F0 CRISPR Screen: Body Size, Adiposity, and Morphology Forest Plots"
subtitle: "Standard Length, Normalised Adiposity, Morphology Effect, Droplet Count, and Droplet Area"
author: "Minchin Lab"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    theme: flatly
    highlight: tango
    fig_width: 6
    fig_height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

# Overview

This document presents forest plots for body size, adiposity, and morphology measures from the F0 CRISPR screen:

1. **Standard Length (SL)** - Body size
2. **Normalised Adiposity (TAT/SL2)** - Size-corrected total adipose tissue  
3. **Morphology Effect (LMM Intercept)** - Deviation in Feret diameter from expected given total adipose
4. **Droplet Count** - Number of lipid droplets per fish
5. **Mean Droplet Area** - Average lipid droplet size

**Statistical approach:** All measures use **Linear Mixed Models (LMM)** with experiment as a random effect to account for batch variation:

- SL, Norm Adiposity, Droplet Count, Droplet Area: `value ~ genotype + (1|experiment)`
- Morphology Effect: `log(Feret) ~ log(AreaSum) + genotype + (1|experiment)`

The morphology effect additionally includes size-correction via `log(AreaSum)` to identify hypertrophic (larger droplets than expected) vs hyperplastic (more, smaller droplets) phenotypes.

**Note:** *lamb2* is excluded (single replicate, n=5 mutants).

# Setup

```{r load-packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(mgcv)
library(lme4)
library(lmerTest)
library(knitr)
library(kableExtra)
```

```{r load-data}
# Load the SL/TAT/measures data
sl_tat_file <- "CRISPR_screen_data_all_measures_v7.csv"
all_data <- read.csv(sl_tat_file, stringsAsFactors = FALSE)

# Load the morphology data (for Feret/morphology effect analysis)
morph_file <- "CRISPR_screen_data_morphology_v7.csv"
morph_data <- read.csv(morph_file, stringsAsFactors = FALSE)

# Standardise experiment names
standardise_exp_name <- function(exp_name) {
  exp_name <- tolower(trimws(exp_name))
  exp_name <- gsub("^exp([0-9]+)$", "expt\\1", exp_name)
  return(exp_name)
}
all_data$experiment <- sapply(all_data$experiment, standardise_exp_name)
morph_data$experiment <- sapply(morph_data$experiment, standardise_exp_name)

cat("SL/TAT data:", nrow(all_data), "observations\n")
cat("Morphology data:", nrow(morph_data), "fish\n")
cat("Genes:", length(unique(morph_data$gene)), "\n")
```

```{r helper-functions}
# Format gene name (italicized for markdown)
format_gene_name <- function(gene, italics = TRUE) {
  if (italics) return(paste0("*", gene, "*"))
  return(gene)
}

# =============================================================================
# LMM-BASED ANALYSIS FOR ALL MEASURES
# =============================================================================
#
# Previously, only morphology used LMM with batch correction. Now ALL measures
# use the same approach for consistency:
#
#   value ~ genotype + (1|experiment)
#
# This accounts for systematic differences between experiments (batch effects)
# and gives more accurate p-values when batch variation exists.
#
# For genes with only 1 experiment, falls back to simple t-test.
# =============================================================================

run_lmm_analysis <- function(data, measure_name) {
  measure_data <- data %>% filter(measure == measure_name)
  
  genes <- unique(measure_data$gene)
  results_list <- list()
  
  for (g in genes) {
    gene_data <- measure_data %>% 
      filter(gene == g, !is.na(value))
    
    n_ctrl <- sum(gene_data$genotype == "control")
    n_mut <- sum(gene_data$genotype == "mutant")
    n_expts <- length(unique(gene_data$experiment))
    
    if (n_ctrl < 3 || n_mut < 3) next
    
    ctrl_mean <- mean(gene_data$value[gene_data$genotype == "control"])
    mut_mean <- mean(gene_data$value[gene_data$genotype == "mutant"])
    
    gene_data$genotype <- factor(gene_data$genotype, levels = c("control", "mutant"))
    
    # Use LMM if >= 2 experiments, otherwise fall back to t-test
    if (n_expts >= 2) {
      tryCatch({
        model <- lmer(value ~ genotype + (1 | experiment), data = gene_data, REML = TRUE)
        coefs <- summary(model)$coefficients
        
        effect_raw <- as.numeric(coefs["genotypemutant", "Estimate"])
        effect_se <- as.numeric(coefs["genotypemutant", "Std. Error"])
        p_value <- as.numeric(coefs["genotypemutant", "Pr(>|t|)"])
        
        # Convert to % change relative to control mean
        effect_pct <- effect_raw / ctrl_mean * 100
        
        results_list[[g]] <- data.frame(
          gene = g,
          n_ctrl = n_ctrl,
          n_mut = n_mut,
          n_expts = n_expts,
          ctrl_mean = ctrl_mean,
          mut_mean = mut_mean,
          effect_pct = effect_pct,
          p_value = p_value,
          method = "LMM",
          stringsAsFactors = FALSE
        )
      }, error = function(e) {
        # If LMM fails, fall back to t-test
        tt <- t.test(gene_data$value[gene_data$genotype == "mutant"],
                     gene_data$value[gene_data$genotype == "control"])
        results_list[[g]] <<- data.frame(
          gene = g,
          n_ctrl = n_ctrl,
          n_mut = n_mut,
          n_expts = n_expts,
          ctrl_mean = ctrl_mean,
          mut_mean = mut_mean,
          effect_pct = (mut_mean - ctrl_mean) / ctrl_mean * 100,
          p_value = tt$p.value,
          method = "t-test (LMM failed)",
          stringsAsFactors = FALSE
        )
      })
    } else {
      # Single experiment - use t-test
      tt <- t.test(gene_data$value[gene_data$genotype == "mutant"],
                   gene_data$value[gene_data$genotype == "control"])
      results_list[[g]] <- data.frame(
        gene = g,
        n_ctrl = n_ctrl,
        n_mut = n_mut,
        n_expts = n_expts,
        ctrl_mean = ctrl_mean,
        mut_mean = mut_mean,
        effect_pct = (mut_mean - ctrl_mean) / ctrl_mean * 100,
        p_value = tt$p.value,
        method = "t-test (single expt)",
        stringsAsFactors = FALSE
      )
    }
  }
  
  results <- do.call(rbind, results_list)
  rownames(results) <- NULL
  
  # BH correction
  results$p_adj <- p.adjust(results$p_value, method = "BH")
  results$sig <- results$p_adj < 0.05
  
  return(results)
}

# Calculate % change from control mean for individual fish (for plotting)
# Centers on control mean so datapoints align with effect estimates
calc_pct_change <- function(data, measure_name) {
  measure_data <- data %>% filter(measure == measure_name)
  
  ctrl_means <- measure_data %>%
    filter(genotype == "control") %>%
    group_by(gene) %>%
    summarise(ctrl_mean = mean(value, na.rm = TRUE), .groups = "drop")
  
  measure_data %>%
    left_join(ctrl_means, by = "gene") %>%
    mutate(pct_change = (value - ctrl_mean) / ctrl_mean * 100)
}
```

```{r forest-plot-function}
# =============================================================================
# FOREST PLOT FUNCTION
# =============================================================================
#
# Creates a forest plot showing three levels of data hierarchy:
#
#   1. SMALL GREY DOTS = Individual mutant fish
#      - Each dot is one fish's % change from control mean
#      - Jittered vertically for visibility
#      - Shows the full distribution of individual variation
#
#   2. MEDIUM GREY CIRCLES = Experiment means
#      - Mean of mutant fish within each experiment/replicate
#      - Shows consistency across biological replicates
#      - Should bracket the overall gene mean if data is consistent
#
#   3. BLACK DIAMOND = Gene mean (overall effect)
#      - For simple measures (SL, adiposity): mean % change vs controls
#      - For morphology: LMM intercept coefficient (batch-corrected)
#      - The primary statistic reported in tables
#
# Visual interpretation:
#   - If circles cluster tightly around diamond = consistent across replicates
#   - If circles spread widely = high batch-to-batch variation
#   - If dots spread widely = high fish-to-fish variation within batches
# =============================================================================

create_forest_plot <- function(plot_data, expt_means, gene_stats, 
                               title, xlab, gene_order) {
  
  # Set factor levels for ordering (genes ordered by effect size)
  plot_data$gene <- factor(plot_data$gene, levels = gene_order)
  expt_means$gene <- factor(expt_means$gene, levels = gene_order)
  gene_stats$gene <- factor(gene_stats$gene, levels = gene_order)
  
  # Create plot - ALL GREY (no colour coding by significance)
 p <- ggplot() +
    # Reference line at 0 (no effect)
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.5) +
    
    # Layer 1: Individual fish (mutants only, small grey dots)
    geom_jitter(
      data = plot_data %>% filter(genotype == "mutant"),
      aes(x = pct_change, y = gene),
      color = "gray70",
      width = 0, height = 0.2, size = 1.2, alpha = 0.35
    ) +
    
    # Layer 2: Experiment means (medium grey circles with border)
    geom_point(
      data = expt_means,
      aes(x = expt_mean, y = gene),
      fill = "gray50", color = "gray30",
      shape = 21, size = 3, stroke = 0.7, alpha = 0.85,
      position = position_jitter(width = 0, height = 0.12, seed = 42)
    ) +
    
    # Layer 3: Gene means (large black diamonds)
    geom_point(
      data = gene_stats,
      aes(x = effect_pct, y = gene),
      shape = 18, size = 4.5, color = "black"
    ) +
    
    # Labels
    labs(
      title = title,
      subtitle = "Dots = fish, Circles = experiment means, Diamonds = gene mean",
      x = xlab,
      y = NULL
    ) +
    
    # Theme
    theme_bw(base_size = 11) +
    theme(
      axis.text.y = element_text(face = "italic", size = 10, color = "gray30"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 12),
      plot.subtitle = element_text(size = 8, color = "gray40"),
      legend.position = "none",
      plot.margin = margin(5, 10, 5, 5)
    )
  
  return(p)
}
```

---

# Analysis 1: Standard Length

```{r sl-analysis}
# =============================================================================
# LMM: SL ~ genotype + (1|experiment)
# Batch-corrected comparison of body size between mutants and controls
# =============================================================================

sl_stats <- run_lmm_analysis(all_data, "SL_um") %>% arrange(effect_pct)
gene_order_sl <- sl_stats$gene

# Individual % change (centered on control mean for plotting)
sl_plot_data <- calc_pct_change(all_data, "SL_um")

# Experiment means (mutants only)
sl_expt_means <- sl_plot_data %>%
  filter(genotype == "mutant") %>%
  group_by(gene, experiment) %>%
  summarise(expt_mean = mean(pct_change, na.rm = TRUE), .groups = "drop")

cat("SL analysis: LMM used for", sum(sl_stats$method == "LMM"), "genes,",
    "t-test for", sum(sl_stats$method != "LMM"), "genes\n")
```

## Statistical Summary

```{r sl-table}
sl_display <- sl_stats %>%
  mutate(
    Gene = sapply(gene, format_gene_name),
    Effect = sprintf("%+.1f%%", effect_pct),
    p_raw = ifelse(p_value < 0.001, sprintf("%.2e", p_value), sprintf("%.4f", p_value)),
    p_adj_fmt = ifelse(p_adj < 0.001, sprintf("%.2e", p_adj), sprintf("%.4f", p_adj)),
    Sig = ifelse(sig, "***", "")
  ) %>%
  select(Gene, n_ctrl, n_mut, n_expts, Effect, p_raw, p_adj_fmt, Sig)

kable(sl_display, 
      col.names = c("Gene", "n ctrl", "n mut", "Expts", "Effect (%)", "p (raw)", "p (BH adj)", ""),
      caption = "Standard Length: LMM with batch correction (value ~ genotype + (1|experiment))",
      escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Forest Plot

```{r sl-forest, fig.width=4, fig.height=10}
p_sl <- create_forest_plot(
  plot_data = sl_plot_data,
  expt_means = sl_expt_means,
  gene_stats = sl_stats,
  title = "Standard Length",
  xlab = "Effect on SL (% change from control)",
  gene_order = gene_order_sl
)
print(p_sl)

# Save
ggsave("forest_plot_SL.svg", p_sl, width = 4, height = 10)
ggsave("forest_plot_SL.png", p_sl, width = 4, height = 10, dpi = 200)
```

---

# Analysis 2: Normalised Adiposity

```{r norm-analysis}
# =============================================================================
# LMM: norm_adiposity ~ genotype + (1|experiment)
# Batch-corrected comparison of size-normalised fat content
# =============================================================================

norm_stats <- run_lmm_analysis(all_data, "norm_adiposity") %>% arrange(effect_pct)
gene_order_norm <- norm_stats$gene

# Individual % change (centered on control mean for plotting)
norm_plot_data <- calc_pct_change(all_data, "norm_adiposity")

# Experiment means (mutants only)
norm_expt_means <- norm_plot_data %>%
  filter(genotype == "mutant") %>%
  group_by(gene, experiment) %>%
  summarise(expt_mean = mean(pct_change, na.rm = TRUE), .groups = "drop")

cat("Norm adiposity analysis: LMM used for", sum(norm_stats$method == "LMM"), "genes,",
    "t-test for", sum(norm_stats$method != "LMM"), "genes\n")
```

## Statistical Summary

```{r norm-table}
norm_display <- norm_stats %>%
  mutate(
    Gene = sapply(gene, format_gene_name),
    Effect = sprintf("%+.1f%%", effect_pct),
    p_raw = ifelse(p_value < 0.001, sprintf("%.2e", p_value), sprintf("%.4f", p_value)),
    p_adj_fmt = ifelse(p_adj < 0.001, sprintf("%.2e", p_adj), sprintf("%.4f", p_adj)),
    Sig = ifelse(sig, "***", "")
  ) %>%
  select(Gene, n_ctrl, n_mut, n_expts, Effect, p_raw, p_adj_fmt, Sig)

kable(norm_display, 
      col.names = c("Gene", "n ctrl", "n mut", "Expts", "Effect (%)", "p (raw)", "p (BH adj)", ""),
      caption = "Normalised Adiposity (TAT/SL2): LMM with batch correction",
      escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Forest Plot

```{r norm-forest, fig.width=4, fig.height=10}
p_norm <- create_forest_plot(
  plot_data = norm_plot_data,
  expt_means = norm_expt_means,
  gene_stats = norm_stats,
  title = "Normalised Adiposity (TAT/SL2)",
  xlab = "Effect on TAT/SL2 (% change from control)",
  gene_order = gene_order_norm
)
print(p_norm)

# Save
ggsave("forest_plot_norm_adiposity.svg", p_norm, width = 4, height = 10)
ggsave("forest_plot_norm_adiposity.png", p_norm, width = 4, height = 10, dpi = 200)
```

---

# Analysis 3: Morphology Effect (LMM Intercept)

This is the key morphology metric from the main analysis: the LMM intercept effect representing uniform shift in Feret diameter after accounting for total adipose area and batch effects.

**Model:** `log(Feret) ~ log(AreaSum) + genotype + (1|experiment)`

This matches the v7 morphology analysis exactly.

```{r morph-lmm-analysis}
# =============================================================================
# LINEAR MIXED MODEL ANALYSIS FOR MORPHOLOGY EFFECT
# =============================================================================
#
# This analysis matches the v7 morphology Rmd exactly.
#
# MODEL: log(Feret) ~ log(AreaSum) + genotype + (1|experiment)
#
# Components:
#   - log(Feret): Mean Feret diameter per fish (log-transformed)
#   - log(AreaSum): Total adipose area per fish (log-transformed) - controls for size
#   - genotype: Fixed effect (control vs mutant) - THE EFFECT WE WANT
#   - (1|experiment): Random intercept for experiment - batch correction
#
# The genotype coefficient represents the UNIFORM SHIFT in log(Feret) for mutants
# vs controls, after accounting for:
#   1. Fish size (via log(AreaSum))
#   2. Batch differences between experiments (random effect)
#
# Positive effect = Hypertrophic (larger droplets than expected for total fat)
# Negative effect = Hyperplastic (smaller droplets than expected for total fat)
#
# The effect is converted to % change: (exp(coefficient) - 1) * 100
# =============================================================================

genes <- unique(morph_data$gene)
lmm_results_list <- list()

for (g in genes) {
  gene_data <- morph_data %>% 
    filter(gene == g, !is.na(mean_feret), !is.na(total_adipose_area),
           total_adipose_area > 0, mean_feret > 0)
  
  if (nrow(gene_data) < 10) next
  
  # Need at least 2 experiments for random effect
  if (length(unique(gene_data$experiment)) < 2) next
  
  # Log transform
  gene_data$log_Feret <- log(gene_data$mean_feret)
  gene_data$log_AreaSum <- log(gene_data$total_adipose_area)
  gene_data$genotype <- factor(gene_data$genotype, levels = c("control", "mutant"))
  
  tryCatch({
    # Fit LMM (matching v7 exactly)
    model <- lmer(log_Feret ~ log_AreaSum + genotype + (1 | experiment), 
                  data = gene_data, REML = TRUE)
    
    coefs <- summary(model)$coefficients
    
    # Extract the genotype effect (mutant vs control)
    intercept_effect <- as.numeric(coefs["genotypemutant", "Estimate"])
    intercept_se <- as.numeric(coefs["genotypemutant", "Std. Error"])
    intercept_t <- as.numeric(coefs["genotypemutant", "t value"])
    intercept_p <- as.numeric(coefs["genotypemutant", "Pr(>|t|)"])
    
    # Convert log-scale effect to percentage
    # e.g., coefficient of 0.15 -> (exp(0.15) - 1) * 100 = +16.2%
    intercept_pct <- (exp(intercept_effect) - 1) * 100
    
    lmm_results_list[[g]] <- data.frame(
      gene = g,
      n_ctrl = sum(gene_data$genotype == "control"),
      n_mut = sum(gene_data$genotype == "mutant"),
      n_experiments = length(unique(gene_data$experiment)),
      intercept_effect = intercept_effect,
      intercept_pct = intercept_pct,
      intercept_se = intercept_se,
      intercept_t = intercept_t,
      intercept_p = intercept_p,
      stringsAsFactors = FALSE
    )
  }, error = function(e) NULL)
}

morph_lmm_results <- do.call(rbind, lmm_results_list)
rownames(morph_lmm_results) <- NULL

# BH correction for multiple testing
morph_lmm_results$p_adj <- p.adjust(morph_lmm_results$intercept_p, method = "BH")
morph_lmm_results$sig <- morph_lmm_results$p_adj < 0.05
morph_lmm_results$direction <- ifelse(morph_lmm_results$intercept_pct > 0, "Hypertrophic", "Hyperplastic")

cat("LMM analysis completed for", nrow(morph_lmm_results), "genes\n")
```

```{r morph-residuals-for-plot}
# =============================================================================
# CALCULATING INDIVIDUAL FISH RESIDUALS FOR FOREST PLOT VISUALISATION
# =============================================================================
#
# The forest plot shows three levels of data:
#   1. Small dots   = individual mutant fish (% deviation from expected Feret)
#   2. Circles      = experiment means (mean of mutant fish within each experiment)
#   3. Diamond      = gene effect (LMM intercept coefficient, batch-corrected)
#
# IMPORTANT: The LMM effect (diamond) represents the genotype coefficient from:
#   log(Feret) ~ log(AreaSum) + genotype + (1|experiment)
# This is the difference between mutant and control AFTER accounting for:
#   - The size-scaling relationship (log(AreaSum))
#   - Batch effects between experiments (random intercept)
#
# For the individual fish and experiment means to align with the LMM effect,
# we need to:
#   1. Calculate residuals from a size-correction model fitted to controls
#   2. Convert to % change: (exp(residual) - 1) * 100
#   3. CENTER on the control mean (subtract control mean from all values)
#
# Without step 3, the control mean would be non-zero (due to regression to mean),
# and the mutant datapoints would be offset from the LMM effect diamond.
# =============================================================================

morph_resid_list <- list()

for (g in unique(morph_lmm_results$gene)) {
  gene_data <- morph_data %>% 
    filter(gene == g, !is.na(mean_feret), !is.na(total_adipose_area),
           total_adipose_area > 0, mean_feret > 0)
  
  if (nrow(gene_data) < 10) next
  
  gene_data$log_Feret <- log(gene_data$mean_feret)
  gene_data$log_AreaSum <- log(gene_data$total_adipose_area)
  gene_data$genotype <- factor(gene_data$genotype, levels = c("control", "mutant"))
  
  # Fit model to controls only to get baseline size-scaling relationship
  ctrl_data <- gene_data %>% filter(genotype == "control")
  
  tryCatch({
    # Step 1: Fit simple linear model on controls (pooled across experiments)
    # This captures the expected Feret diameter given total adipose area
    ctrl_model <- lm(log_Feret ~ log_AreaSum, data = ctrl_data)
    
    # Step 2: Calculate residuals for ALL fish (control + mutant)
    # Residual = observed log(Feret) - expected log(Feret) given size
    gene_data$predicted_log_feret <- predict(ctrl_model, newdata = gene_data)
    gene_data$log_residual <- gene_data$log_Feret - gene_data$predicted_log_feret
    
    # Convert log residuals to % change
    # If residual = 0.15, then % change = (exp(0.15) - 1) * 100 = +16.2%
    gene_data$raw_pct_change <- (exp(gene_data$log_residual) - 1) * 100
    
    # Step 3: CENTER ON CONTROL MEAN
    # The control mean won't be exactly 0 due to the nature of residuals
    # Subtracting it ensures:
    #   - Control fish are centered around 0%
    #   - Mutant fish show deviation FROM controls (matching LMM interpretation)
    #   - Experiment means bracket the overall mutant mean correctly
    ctrl_mean_pct <- mean(gene_data$raw_pct_change[gene_data$genotype == "control"])
    gene_data$pct_change <- gene_data$raw_pct_change - ctrl_mean_pct
    
    morph_resid_list[[g]] <- gene_data %>%
      select(gene, genotype, experiment, mean_feret, total_adipose_area, pct_change)
  }, error = function(e) NULL)
}

morph_plot_data <- bind_rows(morph_resid_list)

cat("Residuals calculated for", length(unique(morph_plot_data$gene)), "genes\n")
cat("Total fish:", nrow(morph_plot_data), "\n")
```

```{r morph-stats}
# =============================================================================
# PREPARE STATS FOR PLOTTING
# =============================================================================
#
# NOTE ON LMM vs SIMPLE MEAN:
# The diamond shows the LMM effect (batch-corrected), while the experiment
# circles show simple means within each experiment. These may differ slightly
# because the LMM weights experiments by precision and accounts for batch effects.
# The individual fish dots and experiment circles should be internally consistent
# (circles bracket the overall simple mean), while the diamond may be slightly
# offset due to batch correction.
# =============================================================================

morph_stats <- morph_lmm_results %>%
  select(gene, n_ctrl, n_mut, effect_pct = intercept_pct, p_value = intercept_p, p_adj, sig, direction) %>%
  arrange(effect_pct)

gene_order_morph <- morph_stats$gene

# Experiment means (mutants only) - these are SIMPLE means, not LMM-adjusted
morph_expt_means <- morph_plot_data %>%
  filter(genotype == "mutant") %>%
  group_by(gene, experiment) %>%
  summarise(expt_mean = mean(pct_change, na.rm = TRUE), .groups = "drop")
```

## Statistical Summary

```{r morph-table}
morph_display <- morph_stats %>%
  mutate(
    Gene = sapply(gene, format_gene_name),
    Effect = sprintf("%+.2f%%", effect_pct),
    p_raw = ifelse(p_value < 0.001, sprintf("%.2e", p_value), sprintf("%.4f", p_value)),
    p_adj_fmt = ifelse(p_adj < 0.001, sprintf("%.2e", p_adj), sprintf("%.4f", p_adj)),
    Sig = ifelse(sig, "***", "")
  ) %>%
  select(Gene, n_ctrl, n_mut, Effect, direction, p_raw, p_adj_fmt, Sig)

kable(morph_display, 
      col.names = c("Gene", "n ctrl", "n mut", "Effect (%)", "Direction", "p (raw)", "p (BH adj)", ""),
      caption = "Morphology Effect (LMM Intercept): Uniform shift in droplet size (batch-corrected)",
      escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Forest Plot

```{r morph-forest, fig.width=4, fig.height=10}
p_morph <- create_forest_plot(
  plot_data = morph_plot_data,
  expt_means = morph_expt_means,
  gene_stats = morph_stats,
  title = "Morphology Effect (LMM Intercept)",
  xlab = "Effect (% change in Feret diameter)",
  gene_order = gene_order_morph
) + 
  coord_cartesian(xlim = c(-100, 100))  # Restrict x-axis to ±100%

print(p_morph)

# Save
ggsave("forest_plot_morphology_effect.svg", p_morph, width = 4, height = 10)
ggsave("forest_plot_morphology_effect.png", p_morph, width = 4, height = 10, dpi = 200)
```

---

# Analysis 4: Droplet Count

```{r droplet-n-analysis}
# =============================================================================
# LMM: droplet_count ~ genotype + (1|experiment)
# Batch-corrected comparison of lipid droplet number
# Note: NOT size-corrected (larger fish may have more droplets)
# =============================================================================

droplet_n_stats <- run_lmm_analysis(all_data, "droplet_count") %>% arrange(effect_pct)
gene_order_droplet_n <- droplet_n_stats$gene

# Individual % change (centered on control mean for plotting)
droplet_n_plot_data <- calc_pct_change(all_data, "droplet_count")

# Experiment means (mutants only)
droplet_n_expt_means <- droplet_n_plot_data %>%
  filter(genotype == "mutant") %>%
  group_by(gene, experiment) %>%
  summarise(expt_mean = mean(pct_change, na.rm = TRUE), .groups = "drop")

cat("Droplet count analysis: LMM used for", sum(droplet_n_stats$method == "LMM"), "genes,",
    "t-test for", sum(droplet_n_stats$method != "LMM"), "genes\n")
```

## Statistical Summary

```{r droplet-n-table}
droplet_n_display <- droplet_n_stats %>%
  mutate(
    Gene = sapply(gene, format_gene_name),
    Effect = sprintf("%+.1f%%", effect_pct),
    p_raw = ifelse(p_value < 0.001, sprintf("%.2e", p_value), sprintf("%.4f", p_value)),
    p_adj_fmt = ifelse(p_adj < 0.001, sprintf("%.2e", p_adj), sprintf("%.4f", p_adj)),
    Sig = ifelse(sig, "***", "")
  ) %>%
  select(Gene, n_ctrl, n_mut, n_expts, Effect, p_raw, p_adj_fmt, Sig)

kable(droplet_n_display, 
      col.names = c("Gene", "n ctrl", "n mut", "Expts", "Effect (%)", "p (raw)", "p (BH adj)", ""),
      caption = "Droplet Count: LMM with batch correction",
      escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Forest Plot

```{r droplet-n-forest, fig.width=4, fig.height=10}
p_droplet_n <- create_forest_plot(
  plot_data = droplet_n_plot_data,
  expt_means = droplet_n_expt_means,
  gene_stats = droplet_n_stats,
  title = "Droplet Count",
  xlab = "Effect on droplet number (% change from control)",
  gene_order = gene_order_droplet_n
)
print(p_droplet_n)

# Save
ggsave("forest_plot_droplet_count.svg", p_droplet_n, width = 4, height = 10)
ggsave("forest_plot_droplet_count.png", p_droplet_n, width = 4, height = 10, dpi = 200)
```

---

# Analysis 5: Mean Droplet Area

```{r droplet-area-analysis}
# =============================================================================
# LMM: mean_droplet_area ~ genotype + (1|experiment)
# Batch-corrected comparison of lipid droplet size
# Note: NOT size-corrected (larger fish may have larger droplets)
# For size-corrected morphology, see Analysis 3 (LMM Intercept)
# =============================================================================

droplet_area_stats <- run_lmm_analysis(all_data, "mean_droplet_area") %>% arrange(effect_pct)
gene_order_droplet_area <- droplet_area_stats$gene

# Individual % change (centered on control mean for plotting)
droplet_area_plot_data <- calc_pct_change(all_data, "mean_droplet_area")

# Experiment means (mutants only)
droplet_area_expt_means <- droplet_area_plot_data %>%
  filter(genotype == "mutant") %>%
  group_by(gene, experiment) %>%
  summarise(expt_mean = mean(pct_change, na.rm = TRUE), .groups = "drop")

cat("Droplet area analysis: LMM used for", sum(droplet_area_stats$method == "LMM"), "genes,",
    "t-test for", sum(droplet_area_stats$method != "LMM"), "genes\n")
```

## Statistical Summary

```{r droplet-area-table}
droplet_area_display <- droplet_area_stats %>%
  mutate(
    Gene = sapply(gene, format_gene_name),
    Effect = sprintf("%+.1f%%", effect_pct),
    p_raw = ifelse(p_value < 0.001, sprintf("%.2e", p_value), sprintf("%.4f", p_value)),
    p_adj_fmt = ifelse(p_adj < 0.001, sprintf("%.2e", p_adj), sprintf("%.4f", p_adj)),
    Sig = ifelse(sig, "***", "")
  ) %>%
  select(Gene, n_ctrl, n_mut, n_expts, Effect, p_raw, p_adj_fmt, Sig)

kable(droplet_area_display, 
      col.names = c("Gene", "n ctrl", "n mut", "Expts", "Effect (%)", "p (raw)", "p (BH adj)", ""),
      caption = "Mean Droplet Area: LMM with batch correction",
      escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Forest Plot

```{r droplet-area-forest, fig.width=4, fig.height=10}
p_droplet_area <- create_forest_plot(
  plot_data = droplet_area_plot_data,
  expt_means = droplet_area_expt_means,
  gene_stats = droplet_area_stats,
  title = "Mean Droplet Area",
  xlab = "Effect on droplet area (% change from control)",
  gene_order = gene_order_droplet_area
)
print(p_droplet_area)

# Save
ggsave("forest_plot_droplet_area.svg", p_droplet_area, width = 4, height = 10)
ggsave("forest_plot_droplet_area.png", p_droplet_area, width = 4, height = 10, dpi = 200)
```

---

# Summary

```{r summary-table}
# Key genes
key_genes <- c("mmp14b", "foxp1b", "txnipa", "kazna", "tbx15")

# Merge all stats for key genes
summary_df <- data.frame(gene = key_genes) %>%
  left_join(sl_stats %>% select(gene, SL_pct = effect_pct, SL_sig = sig), by = "gene") %>%
  left_join(norm_stats %>% select(gene, Norm_pct = effect_pct, Norm_sig = sig), by = "gene") %>%
  left_join(morph_stats %>% select(gene, Morph_pct = effect_pct, Morph_sig = sig, Morph_dir = direction), by = "gene") %>%
  left_join(droplet_n_stats %>% select(gene, Drop_n_pct = effect_pct, Drop_n_sig = sig), by = "gene") %>%
  left_join(droplet_area_stats %>% select(gene, Drop_area_pct = effect_pct, Drop_area_sig = sig), by = "gene")

summary_display <- summary_df %>%
  mutate(
    Gene = paste0("*", gene, "*"),
    SL = sprintf("%+.1f%s", SL_pct, ifelse(SL_sig, "***", "")),
    `Norm Adip` = sprintf("%+.1f%s", Norm_pct, ifelse(Norm_sig, "***", "")),
    `Morph Effect` = sprintf("%+.1f%% (%s)%s", Morph_pct, substr(Morph_dir, 1, 5), ifelse(Morph_sig, "***", "")),
    `Drop Count` = sprintf("%+.1f%s", Drop_n_pct, ifelse(Drop_n_sig, "***", "")),
    `Drop Area` = sprintf("%+.1f%s", Drop_area_pct, ifelse(Drop_area_sig, "***", ""))
  ) %>%
  select(Gene, SL, `Norm Adip`, `Morph Effect`, `Drop Count`, `Drop Area`)

kable(summary_display,
      caption = "Summary: Key Gene Phenotypes (*** = BH-adjusted p < 0.05)",
      escape = FALSE,
      align = c("l", "r", "r", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Interpretation

```{r interpretation, results='asis'}
cat("
### *mmp14b*
- **SL:** +2% (ns) - Normal body size
- **Normalised adiposity:** -1% (ns) - Normal total fat
- **Morphology effect:** Hypertrophic (larger droplets than expected)
- **Droplet count:** +2% (ns) - Normal number
- **Droplet area:** +52% - Larger droplets

Hypertrophic phenotype: Same total fat in larger droplets.

---

### *foxp1b*
- **SL:** +1% (ns) - Normal body size
- **Normalised adiposity:** -13% (trending) - Consistent with stable mutant phenotype
- **Morphology effect:** Hypertrophic
- **Droplet count:** +18% (ns)
- **Droplet area:** +73%

F0 data now consistent with stable mutants after control batch matching.

---

### *txnipa*
- **SL:** +9% - Larger fish
- **Normalised adiposity:** +25% - More fat (validates mouse *Txnip* KO)
- **Morphology effect:** Hypertrophic
- **Droplet count:** +24% (ns)
- **Droplet area:** +67%

True adiposity phenotype with larger droplets.

---

### *kazna*
- **SL:** +10% - Larger fish
- **Normalised adiposity:** +60%*** - **Significantly more fat**
- **Morphology effect:** Hypertrophic
- **Droplet count:** +90%*** - **Many more droplets**
- **Droplet area:** +56%

Strong adiposity phenotype. **Novel finding** - no published adipose phenotype.

---

### *tbx15*
- **SL:** Normal body size
- **Normalised adiposity:** Significantly reduced fat
- **Morphology effect:** Hyperplastic (smaller droplets than expected)
- **Droplet count:** Reduced
- **Droplet area:** Smaller droplets

**Significant adiposity phenotype** - reduced fat with hyperplastic morphology. Validates known role in adipose development.
")
```

---

# Excluded Gene: *lamb2*

*lamb2* was excluded from analyses:

- Single experiment only
- n = 5 mutants (vs 18 controls)

**Preliminary results:**

- SL: +35%*** (larger fish)
- Normalised adiposity: +127%** (very large increase)

Requires independent validation.

---

# Export Results

```{r export}
# =============================================================================
# EXPORT STATISTICS
# All measures use LMM with batch correction: value ~ genotype + (1|experiment)
# Morphology additionally includes size-correction via log(AreaSum)
# =============================================================================

# Combine all stats (note: columns may differ slightly between measures)
all_stats <- bind_rows(
  sl_stats %>% mutate(measure = "SL_um"),
  norm_stats %>% mutate(measure = "norm_adiposity"),
  morph_stats %>% mutate(measure = "morphology_LMM_intercept"),
  droplet_n_stats %>% mutate(measure = "droplet_count"),
  droplet_area_stats %>% mutate(measure = "mean_droplet_area")
)

# Save combined statistics
write.csv(all_stats, "CRISPR_screen_all_measures_statistics.csv", row.names = FALSE)

# Also save the full LMM results for morphology (includes additional columns)
write.csv(morph_lmm_results, "CRISPR_screen_morphology_LMM_results.csv", row.names = FALSE)

cat("Statistical approach: LMM with batch correction for all measures\n")
cat("Model: value ~ genotype + (1|experiment)\n")
cat("Morphology model: log(Feret) ~ log(AreaSum) + genotype + (1|experiment)\n\n")
cat("Results exported:\n")
cat("- CRISPR_screen_all_measures_statistics.csv\n")
cat("- CRISPR_screen_morphology_LMM_results.csv\n")
cat("\nForest plots saved as SVG and PNG (3\" x 10\"):\n")
cat("- forest_plot_SL.svg/png\n")
cat("- forest_plot_norm_adiposity.svg/png\n")
cat("- forest_plot_morphology_effect.svg/png\n")
cat("- forest_plot_droplet_count.svg/png\n")
cat("- forest_plot_droplet_area.svg/png\n")
```

---

# Session Info

```{r session-info}
sessionInfo()
```

---

*Analysis completed: `r Sys.time()`*
